---
title: "benchmark of linear solver"
author: "Yinan Zhu"
date: "September 28, 2017"
output:
  pdf_document: default
  html_document: default
---
```{r}
library(microbenchmark)
source('~/GitHub/SDS385-course-work/Excercise 1/linear equation solver/linear solver.R')
```
First, consider two solvers, of which "linearsolver" use inverse matrix and "linearsolverlu" use LU decomposition, we should check that they give the correct result, using data generated from a linear model with gaussian noise

```{r}
n=3000
p=20
beta=matrix(1,nrow=p,ncol=1)
X=matrix(runif(n*p,0,1),nrow=n,ncol=p)
y=X%*%beta+rnorm(n,sd=1)
```
```{r}
print(linearsolve(y,X))
```
```{r}
print(linearsolvelu(y,X))
```

Now we fix n, the number of points to be constant, and varies p, the number of features. Test the time it takes to solve

```{r}
t=as.vector(rep(0,10))
tlu=t
for(i in 1:10){
p=200*i
beta=matrix(1,nrow=p,ncol=1)
X=matrix(runif(n*p,0,1),nrow=n,ncol=p)
y=X%*%beta+rnorm(n,sd=1)

l=microbenchmark(linearsolve(y,X),times=1L)
t[i]=l$time/1000000000
l=microbenchmark(linearsolvelu(y,X),times=1L)
tlu[i]=l$time/1000000000
}
```
```{r}
plot(t~seq(from=200,to=2000,by=200),xlab='p')
points(tlu~seq(from=200,to=2000,by=200),col='red')
```

Now we add "linearsolvesparse" which use the sparse matrix class

```{r}
p=20
mask=matrix(rbinom(n*p,1,0.05),nrow=n,ncol=p)
X=matrix(runif(n*p,0,1),nrow=n,ncol=p)
beta=matrix(1,nrow=p,ncol=1)
Xsparse=X*mask
y=Xsparse%*%beta+rnorm(n,sd=10)
```
```{r}
print(linearsolve(y,Xsparse))
```
```{r}
print(linearsolvelu(y,Xsparse))
```
```{r}
print(linearsolvesparse(y,Xsparse))
```

To benchmark the performance, we keep the size of equation constant and varies the density of observed data

```{r}
t=as.vector(rep(0,6))
tlu=t
ts=t
p=1600
X=matrix(runif(n*p,0,1),nrow=n,ncol=p)
beta=matrix(1,nrow=p,ncol=1)
for(i in 1:6){
  mask=matrix(rbinom(n*p,1,0.1*i),nrow=n,ncol=p)
  Xsparse=X*mask
  y=Xsparse%*%beta+rnorm(n,sd=10)
  l=microbenchmark(linearsolve(y,Xsparse),times=1L)
  t[i]=l$time/1000000000
  l=microbenchmark(linearsolvelu(y,Xsparse),times=1L)
  tlu[i]=l$time/1000000000
  l=microbenchmark(linearsolvesparse(y,Xsparse),times=1L)
  ts[i]=l$time/1000000000
}
```
```{r}
plot(t~seq(from=0.1,to=0.6,by=0.1),ylim=c(0,15),xlab='density')
points(tlu~seq(from=0.1,to=0.6,by=0.1),col='red')
points(ts~seq(from=0.1,to=0.6,by=0.1),col='blue')
```

